@c -*-texinfo-*-

@chapter Introduction

Back in the '70s and '80s Lisp workstations were an important development.  A
portable version of one of those environments was created at that time called
Medley.  The source code to that original system has been made
open-source, ported, and modernized.  This introduction is about that
system.

Medley originally supported Interlisp (an older dialect of Lisp) but
grew to support Common Lisp as well.  Medley, therefore, is a software
development environment that embodies Interlisp and Common Lisp.

The Medley development environment is very different from the file-based
model used today.  The value of this system is to provide the ability
to learn about, explore, and possibly utilize these benefits in future
projects.

The majority of the system is written in Lisp and rides on a portable
virtual machine written in portable C.  The system uses an X11 window
to display the bitmap in which the Lisp code manages. The system is
very portable and has been ported to Linux, Macintosh, and other
systems.

The purpose of this document is to provide access to this system to
someone who is already familiar with current Common Lisp development
environments in as brief a form as possible.

@section History

Lisp was formulated by John McCarthy in 1958 at MIT.  Lisp quickly
became the language of choice for the burgeoning field of Artificial
Intelligence (AI) research.  Lisp quickly spread in the academic
community.  Lisp also spread in the commercial world to support
research projects done at those institutions.

Due in part to the youth of the industry, institutions often
implemented their own version of Lisp, and those versions almost
always differed from other Lisp implementations in ways that suited
each particular institution.  The two most common implementations were
MacLisp and Interlisp.  MacLisp came from MIT (the East coast), and
Interlisp (which was derived from BBN Lisp) came from the Xerox Palo
Alto Research Center in California (the West coast).  Both systems
bled into the commercial world.

Back in the 1960's through the early 80's, much research in AI was
funded by DARPA which was a government agency.  DARPA became tired of
supporting projects written in incompatible Lisp implementations.
They therefore issued an edict demanding a single, portable version of
Lisp.  Thus became Common Lisp.

While having a single version of Lisp had the benefit of making
software more portable, it also had the negative effect of 
somewhat stifling the exploration in the language.

MacLisp was written in assembler code for particular machines.  Since
those machines don't exist anymore, MacLisp itself has disappeared.
Interlisp was also largely written in assembler and those systems too
have disappeared.  However, there was at least one version of Interlisp
written in C.  That is, our system, Medley.  Because it was written in
C (to some, a portable assembler), it is still available today.

From its inception up until very recently, Medley was a commercial
product and was unavailable freely or in source-code form.  Recently,
the system has been made open-source.

Due to the efforts of some of the original authors of Interlisp and
others in the community, Medley is being made available to the
general public.







